---
title: "Gov 50 Final Project"
author: "Faith Oji"
description: "Race's Role in Romance"
output:
  distill::distill_article:
    self_contained: false
---



## Milestone One

Honestly, I'm not exactly sure what I want my final project to be about quite yet. I have a lot of interests in the lasting effects of slavery on systematic racism in the united states, as well as how family histories (in relation to the Slave Trade) influence political activity today. I'm not yet sure, though, about how I would tie this in to my final project. I'd also be interested in how immigration rates affect different aspects in our system. I also potentially could do a project on the prison systems and abolition. I'm just not sure yet what direction I want to go in or how I'd go in that direction.

## Milestone Two

Upon further research, I've decided to completely change the course of my project. I've decided, instead, to do a project surrounding marriage, specifically pulling from a data set in relation to a study called 'Marriage is Not Dead'. My research question is whether or not race has an impact on tendency towards marriage. My hypothesis is that, throughout the years, Black People in America will have the highest proportion of individuals that have never been married. This hypothesis stems from a few points, with the main one being the effect of systematic racism in the United States. It's common knowledge that marriage is very expensive. However, the systems within this country make it disproportionately harder for Black People to acquire and maintain large amounts of wealth, and as such, they would be less likely to take on the financial burden of a wedding. There is also the factor of the prison systems breaking up Black families, which would also contribute to a lower marriage rate. 

My explanatory variables of interest are 'White', 'Black', and 'Hisp', which represent White People, Black or African-American People, and Hispanic People, respectively. These are all different races. The outcome variable of interest would be the proportion of the population that is unmarried per year. This was measured through the survey data, and the trend could be noted through a line graph. An observed pattern in the data that would support this hypothesis would be whether or not the proportion of unmarried Black People was significantly higher than the other races, as well as the overall proportion of unmarried individuals throughout the years. My hypothesis would be disproved if the proportions were similar amongst all the populations, or if a different race had a significantly higher proportion of unmarried individuals. 

Below is a link to the original dataset. 
https://github.com/fivethirtyeight/data/blob/master/marriage/both_sexes.csv

## Milestone Three

```{r}
library(tidyverse)

both_sexes <- read_csv("both_sexes.csv")

both_sexes
```

```{r}

race_compiled <- both_sexes |>
  select(year, Black_2534, Black_3544, Black_4554, Hisp_2534, Hisp_3544, Hisp_4554, White_2534, White_3544, White_4554)|>
  group_by(year)|>
  mutate(Black = (Black_2534+ Black_3544+ Black_4554)/3,
         Hispanic = (Hisp_2534+ Hisp_3544+ Hisp_4554)/3,
         White = (White_2534+ White_3544+ White_4554)/3,)|>
  select(year, Black, White, Hispanic) 

race_compiled
```

```{r}
library(ggplot2)
race_graph <- race_compiled|>
  ggplot(aes(x = year))+
  geom_line(aes(y= Black, colour = "Black"))+
  geom_line(aes(y= White, colour = "White"))+
  geom_line(aes(y= Hispanic, colour = "Hispanic"))+
  labs(title = "Unmarried Population Over The Years",
       x = "Year",
       y = "Proportion of Unmarried Population",
       colour = "Legend") +
  scale_color_manual(name = "Race", values= c("Black"= "deeppink1", "White"= "deepskyblue1", "Hispanic"= "darkorchid1"))

race_graph
```
The above graph is the current version of my data visualisation. The three lines correspond with each of the races involved in the study (Black, White, and Hispanic). Though the current legend doesn't include accurate labels, the green line corresponds with the Black population, the red line corresponds with the Hispanic population, and the blue line corresponds with the White population. In future versions of the visualisation, I hope to figure out what's causing the lines to be mislabeled in the legend. However, the data shows an interesting trend that seems to prove the hypothesis that that Black populations have far higher proportions of unmarried people within the populations. It also highlights an upward trend in the unmarried proportion across all populations.  


###Milestone Four

```{r}
race_compiled <- both_sexes |>
  select(year, Black_2534, Black_3544, Black_4554, Hisp_2534, Hisp_3544, Hisp_4554, White_2534, White_3544, White_4554)|>
  group_by(year)|>
  mutate(Black = (Black_2534+ Black_3544+ Black_4554)/3,
         Hispanic = (Hisp_2534+ Hisp_3544+ Hisp_4554)/3,
         White = (White_2534+ White_3544+ White_4554)/3,)|>
  select(year, Black, White, Hispanic) 

race_compiled

knitr::kable(race_compiled, digits = 2, col.names = c("Year", "Black Population", "White Population", "Hispanic Population"))

```

This table is a neat visualisation of the numerical proportions of the unmarried populations over the years. Through this table, it is even easier to note how the Black populations have a drastically larger proportion of unmarried individuals over the years, and over generations, which alludes to the lasting effects of inequality across generations. Additionally, the visualisation from Milestone Three has been updated to include proper labels for the different lines.

## Introduction

For this project, I decided to explore data relating to marriage rates in America since the 1960s. I specifically wanted to explore the link between the race of American citizens and the effect that that may have on the proportion of the unmarried people within those subsets of the population. I chose this topic because, to be honest, the idea of marriage has always been one that has interested me, specifically the disparities that I had noticed amongst different groups surrounding the perception of how much of their population had found traditional marriages compared to other types of relationships. I wanted to see if these stereotypes had any statistical backing.

My research question in this case was whether or not race had an impact on the percentage of a population that remained married. If so, which racial group would have the highest proportion of unmarried individuals? My hypothesis was based in the idea that marriage is a luxury, and since its foundations, the systems in America have largely prevented Black Americans from experiencing much luxury. As such, my hypothesis is that there would be a notable difference in the proportion of the unmarried population, with Black Americans having the largest proportion throughout the years.

To support my hypothesis, my visualisations would have to show two things. First, it would have to show a large gap between the proportions of unmarried populations across the three racial groups identified: White, Hispanic, and Black. This is because, without an initial notable difference in the proportions, there would be no need to further analyse the data. Secondly, and more importantly, there would have to be an identifiable difference between the proportion of the Black unmarried population in comparison to those of the other two racial groups. If there were no initial difference, my hypothesis would have to be proven false. Additionally, if a different racial group had a noticeably higher proportion of unmarried individuals, my hypothesis would have to be proven false. 

## Data and Research Method

In this project, I had to modify my data set in order to solely show my variables of interest. Originally, the dataset included additional information, such as the region of the individuals, the level of education, employment history, economic status, and the presence of children within the home. While these are important variables that may have an effect on the marruage rates, for the purpose of my project, I wanted to primarily focus on the effect of race on long-term relationships. Additionally, the original data set separated each racial subgroup into three different subgroups, based on age. Because I'm interested in the effects across the whole population, not just by age group, I had to modify the dataset once again to find the means of each of these separations. I pulled the necessary information out of the originsl dataset, 'both_sexes', and compiled it into my unique dataset, 'race_compiled'

# Original Data Set
```{r}
library(tidyverse)

both_sexes <- read_csv("both_sexes.csv")

both_sexes
```

# Modified Data Set
```{r}

race_compiled <- both_sexes |>
  select(year, Black_2534, Black_3544, Black_4554, Hisp_2534, Hisp_3544, Hisp_4554, White_2534, White_3544, White_4554)|>
  group_by(year)|>
  mutate(Black = (Black_2534+ Black_3544+ Black_4554)/3,
         Hispanic = (Hisp_2534+ Hisp_3544+ Hisp_4554)/3,
         White = (White_2534+ White_3544+ White_4554)/3,)|>
  select(year, Black, White, Hispanic) 

race_compiled
```

As you can see, the modified dataset focuses on the year the data was collected and has three additional colums: "Black", "White", and "Hispanic". I calculated these columns using the original data. For the "Black" column, I found the sum of the proportion of unmarried individuals in each age range (25-34, 35-44, 45-54), and then divided that number by three. I followed the same process for each of the subsequent racial columns as well. 