---
title: "Gov 50 Final Project"
author: "Faith Oji"
description: "Race's Role in Romance"
output:
  distill::distill_article:
    self_contained: false
---
## Introduction

For this project, I decided to explore data relating to marriage rates in America since the 1960s. I specifically wanted to explore the link between the race of American citizens and the effect that that may have on the proportion of the unmarried people within those subsets of the population. I chose this topic because, to be honest, the idea of marriage has always been one that has interested me, specifically the disparities that I had noticed amongst different groups surrounding the perception of how much of their population had found traditional marriages compared to other types of relationships. I wanted to see if these stereotypes had any statistical backing.

My research question in this case was whether or not race had an impact on the percentage of a population that remained married. If so, which racial group would have the highest proportion of unmarried individuals? My hypothesis was based in the idea that marriage is a luxury, and since its foundations, the systems in America have largely prevented Black Americans from experiencing much luxury. As such, my hypothesis is that there would be a notable difference in the proportion of the unmarried population, with Black Americans having the largest proportion throughout the years.

To support my hypothesis, my visualisations would have to show two things. First, it would have to show a large gap between the proportions of unmarried populations across the three racial groups identified: White, Hispanic, and Black. This is because, without an initial notable difference in the proportions, there would be no need to further analyse the data. Secondly, and more importantly, there would have to be an identifiable difference between the proportion of the Black unmarried population in comparison to those of the other two racial groups. If there were no initial difference, my hypothesis would have to be proven false. Additionally, if a different racial group had a noticeably higher proportion of unmarried individuals, my hypothesis would have to be proven false. 

## Data and Research Method

In this project, I had to modify my data set in order to solely show my variables of interest. Originally, the dataset included additional information, such as the region of the individuals, the level of education, employment history, economic status, and the presence of children within the home. While these are important variables that may have an effect on the marriage rates, for the purpose of my project, I wanted to primarily focus on the effect of race on long-term relationships. Additionally, the original data set separated each racial subgroup into three different subgroups, based on age. Because I'm interested in the effects across the whole population, not just by age group, I had to modify the dataset once again to find the means of each of these separations. I pulled the necessary information out of the original dataset, 'both_sexes', and compiled it into my unique dataset, 'race_compiled'. The data present comes from a cross-sectional study. 

# Original Data Set
```{r}
library(tidyverse)

both_sexes <- read_csv("both_sexes.csv")

both_sexes
```

# Modified Data Set
```{r}

race_compiled <- both_sexes |>
  select(year, Black_2534, Black_3544, Black_4554, Hisp_2534, Hisp_3544, Hisp_4554, White_2534, White_3544, White_4554)|>
  group_by(year)|>
  mutate(Black = (Black_2534+ Black_3544+ Black_4554)/3,
         Hispanic = (Hisp_2534+ Hisp_3544+ Hisp_4554)/3,
         White = (White_2534+ White_3544+ White_4554)/3,)|>
  select(year, Black, White, Hispanic) 

race_compiled
```

As you can see, the modified dataset focuses on the year the data was collected and has three additional columms: "Black", "White", and "Hispanic". I calculated these columns using the original data. For the "Black" column, I found the sum of the proportion of unmarried individuals in each age range (25-34, 35-44, 45-54), and then divided that number by three. I followed the same process for each of the subsequent racial columns as well.

# Data Visualisation

```{r}
library(ggplot2)
race_graph <- race_compiled|>
  ggplot(aes(x = year))+
  geom_line(aes(y= Black, colour = "Black"))+
  geom_line(aes(y= White, colour = "White"))+
  geom_line(aes(y= Hispanic, colour = "Hispanic"))+
  labs(title = "Proportion of Unmarried Population Over The Years",
       x = "Year",
       y = "Proportion of Unmarried Population",
       colour = "Legend") +
  scale_color_manual(name = "Race", values= c("Black"= "deeppink1", "White"= "deepskyblue1", "Hispanic"= "darkorchid1"))

race_graph

```
```{r}

knitr::kable(race_compiled, digits = 2, col.names = c("Year", "Black Population", "White Population", "Hispanic Population"))

```

```{r}

black <- race_compiled|>
  select(Black)

black

ggplot(black, aes(x = year, y = Black)) +
geom_point(color = "black") +
labs(
x = "Year",
y = "Unmarried Proportion",
title = "Unmarried Proportion of Black Population Per Year")+
geom_smooth(method = "lm", se = FALSE, color = "deeppink1", size = 1.5)

black_fit <- lm(Black ~ year, data = black)
black_fit
```

```{r}
hispanic <- race_compiled|>
  select(Hispanic)

hispanic

ggplot(hispanic, aes(x = year, y = Hispanic)) +
geom_point(color = "black") +
labs(
x = "Year",
y = "Unmarried Proportion",
title = "Unmarried Proportion of Hispanic Population Per Year")+
geom_smooth(method = "lm", se = FALSE, color = "darkorchid1", size = 1.5)

hispanic_fit <- lm(Hispanic ~ year, data = hispanic)
hispanic_fit
```
```{r}
white <- race_compiled|>
  select(White)

white

ggplot(white, aes(x = year, y = White)) +
geom_point(color = "black") +
labs(
x = "Year",
y = "Unmarried Proportion",
title = "Unmarried Proportion of White Population Per Year")+
geom_smooth(method = "lm", se = FALSE, color = "deepskyblue1", size = 1.5)

white_fit <- lm(White ~ year, data = white)
white_fit
```

The first visualisation shows the comparison of proportions across all of the discussed racial populations, while the following scatterplots concern the individual regressions. 

## Results

The above visualisations show two major aspects. Firstly, both the line graph and the table prove that there is a notable difference between the proportions of unmarried populations. The graph shows the pink line, which corresponds with the Black population, consistently being above the lines corresponding with the other two racial groups. This is further supported by the table, which, in each year, has the highest value in the leftmost column, which corresponds with the Black population.

Furthermore, the graph shows an interesting trend. While the proportions of the Hispanic and White populations' proportions remain relatively close to one another, having a maximum difference of four percentage points in 2009, 2011, and 2012, the difference between those to populations and the Black population grows exponentially over the years. What started as a two percentage point difference in 1960 grew to close to a twenty percentage point difference by 2012. The trend of the graph also suggests that this increase would continue past the scope of the study.

When looking at the regressions, it is even clearer that the proportion of the unmarried Black population is drastically different in comparison to that of the other two populations. The coefficients in each of the broken down scatterplots are representative of how much the proportion of unmarried individuals in the population increased per year. For both the White and Hispanic populations, that number was relatively the same, with both having a coefficient of approximately .003. However, the coefficient when concerning the Black population was over twice that figure, at approximately.007. However, based on the nature of the study, there isn't enough evidence to claim this to be a causal effect. However, there is a strong correlation.

## Conclusion

In conclusion, there is a strong correlation between race and the proportion of the population that is unmarried. When it came to the Black population, a significantly higher proportion of the population was unmarried than in any of the other populations. The line graph was most useful at showing the sheer difference between the population proportions, but the regression lines assisted in clarifying that race also has an effect on the growth of the unmarried proportions in these populations. There are quite a few limitations in this dataset, as it doesn't include all possible age groups of married individuals, doesn't include information about potential confounders, and doesn't include information about other races. A way to improve this study would be to collect that information to form a better understanding of the issue. 